An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-30 draft.tex
Exp 2 images

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously.
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population \cite{li2015multi}. Researchers use the term
multi-population based methods when generally referring to techniques using
many populations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but must maintain synchronism with
the other operations,  all operations need to finish to continue to the next
step. For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. Following the previous example, now the parent operation can
continue to the next iteration, even if a worker has not finished its work.  In
the literature, we can find many asynchronous algorithms
\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time,
and scalability. In the particular case of asynchronous and cloud-native
multi-population algorithms, recent works propose an asynchronous communication
through a central repository \cite{sofea:cec2012, JSON} or message queues
\cite{salza2019speed, guervos2018introducing}, in this work we follow this
practice.

@@

When designing efficient multi-population algorithms, researchers need to
consider additional issues \cite{Ma2019}. These include the number and size of populations,
the interaction between them, the search area of each population, and the search
strategy and parametrization of each population. We must take into account
these high-level issues when designing a parallel architecture. Researchers 
and algorithm designers willing to  adapt to the cloud their
multi-population solutions face the additional challenges:
\begin{itemize}
    \item They need to change their current solutions to a reactive architecture in which processes
    communicate with each other by exchanging messages asynchronously and react
    to a continuous stream of data. One of the options is to consider
    populations as messages to be modified asynchronously by functions.

@@

Taking these factors into account, in this work, we present EvoSwarm, 
a multi-container Docker application that reactively processes isolated and
heterogeneous populations. The platform uses the Docker Compose tool for
defining and running experiments as multi-container applications. 
% Docker compose is not really for production, more for development environments, 
% you have to justify this. Also, it does not handle scaling so need to justify it too - JJ
% justification: - Mario
We preferred the use of a container-based application for this prototype because
it is more suitable for development and replication without the need for
additional components. Nevertheless, the application is compatible with other
orchestration and deployment technologies if more scalability is needed. 
With additional configuration settings, applications can be deployed 
to a Docker Swarm or a Kubernetes system.

Docker Container images host the stateless functions responsible for running 
a search strategy on several populations, these functions could also be ported
directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
% Need to improve the next paragraph or maybe move it to another place - Mario
Researchers then define the configuration of services used by
a multi-population algorithm using a YAML file. Any researcher can deploy and
start all the services from this configuration file with a single command. Once
the services are up, multiple instances of an experiment can be sent to a queue
for their execution. The platform includes containerized services that are
compatible with other container orchestration technologies like Kubernetes and
Docker Swarm. The services included are a Redis Server used for the Message
Queue implementation and logging and an Experiment Controller implemented in
Python. Developers of population processing functions have the freedom to
implement new population-based algorithms or operators in the language they
choose inside a container, as they only need to subscribe to a channel of the
Message Queue and have the ability to read the population and parameters as JSON  
file.

@@

% Also, multipopulation design of what?x
Multi-population based methods divide the original population into
smaller populations or islands, with every population carrying out the
algorithm independently, with synchronous or asynchronous communication with the
rest of the islands \cite{Ma2019}.% exemplify how this is done in different
                     % metaheuristics, especially those we will be
                     % using here - JJ
This relative isolation helps in maintaining an overall
diversity since each population will search in a particular area, at least
between communications. The recombination mentioned above (mixing) or migration
between populations is needed to avoid a premature convergence of candidate
solutions since smaller populations are known to perform better for a given
problem than bigger populations \cite{li2016multi,wu2016differential}. % Smaller populations converge
                                % prematurely or perform better? - JJ
However, it gives them the added advantage of
parallel operation. Additionally, and in some cases, multi-population algorithms
scale better than expected due to the interaction between the algorithm and the
parallelism of the operation \cite{ALBA20027}. % Need to clarify this
                                % and if it extends to Swarm
                                % algorithms too
                                % Also, is your objective to achieve this superlinear speedup? - JJ

However, in most cases, algorithms applied to each population are
homogeneous, or at any rate, the same variant of the algorithm. As long as this
parallel operation is not synchronous, other population-based algorithms, or, as
a matter of fact, any algorithm, could be easily integrated. That is why several
works based on multi-population are heterogeneous, integrating various
optimization algorithms, and often performing better than single-population or
homogeneous optimization algorithms \cite{wu2016differential,nseef2016adaptive}.

Heterogeneous algorithms add another degree of freedom to the problem of finding
the correct parameter settings for an algorithm; because some parameters affect
the accuracy of the solution and the convergence speed of the algorithms as they
tip the balance between exploration and exploitation of the search space. On the
other hand, current studies show that by having a high number of populations
interacting in parallel, the effect of the individual parameters of each
population is compensated by those selected in other
populations \cite{li2016multi,tanabe2013evaluation}. % reference - JJ

In this work, we will use random settings within a specific range as results have shown
this is a valid solution to this problem \cite{garcia2014randomized}. % This should have been
                                % mentioned in the introduction - JJ

Some parameters, specially the population size, are
kept fixed in order to control more easily the execution of the algorithm. For
instance, by having the size of populations fixed, it is easier to control
the number of evaluations and the communication costs, when the algorithm is in
operation. % And to keep them more or less synchronized, which can be
           % interesting, too... But this should go to the
           % introduction also - JJ

Combining multiple algorithms, with different parameters, interacting with each
other at the same time, can benefit from the strengths of each. For instance, a
genetic algorithm could find a promising global solution that is not optimal
while another algorithm, more suitable for a local search, finds the global
optimum. This approach has been followed extensively in recent years with
success \cite{li2015multi,godio2016multi,biswas2014co}. % References - JJ
Moreover, there is a need for frameworks, architecture, and
implementation models that can allow researchers the development of new
parallel, asynchronous, heterogeneous, and parameter-free algorithms in a scalable way.

@@

The most elemental data structure used throughout the system is the population; and at a higher
level, we can view the system as a continuous stream of populations
flowing from one process to the other. By "continuous stream," we mean that ideally,
everything must happen asynchronously without components needing to wait for
others. In this kind of system, this streaming functionality is normally
implemented by using a message queue system. For instance, if one process needs
to communicate a population to another, it must pack the population into
a message, and then send the message 
to the queuing system without waiting for a response; this
means that after the process pushes the message, it continues its execution
without waiting for a result. On the other end, the recipient subscribes to the
message queue by defining an event handler method that is going to be triggered when a
message is pushed to it. The two components at each end of the queue are
Producers that push messages to the queue and Consumers that pull them. This
pattern is an essential component of a reactive architecture and makes
it 
highly scalable; one of the reasons for this is the use of functions
that have no secondary effects or access global storage, that is,
stateless functions. These functions do not need to read, keep, or modify data outside of the method. If
Consumers are implemented using stateless functions, and thus, there is no difference
between having one or many copies of the same function pulling work from the queue all at the
same time, because no side effects, including unwanted resource
locking problems,  are going to result from this concurrency.
% This is important and we should probably mention that in the
% introduction - JJ 
Based on these general principles we define next the proposed model.

@@

\subsection{Event Driven Model} 
\label{edm}
Based on a reactive architecture we proposed in a previous work
\cite{guervos2018introducing}, we now describe the general architecture shown in
Figure \ref{fig:kafkEO}. Again, we can explain the model using the analogy
of producers and consumers of messages. First, we can see that
there are two queues, one labeled \texttt{Input}, and the other one \texttt{Output}. In the diagram, 
push operations on a queue are represented by solid arrows connecting to the left side
of the queue box, and pull or pop operations as solid arrows leaving from the right side.
Also, the architecture has at least four processes indicated in the diagram as
swimlanes: First, the \texttt{Setup} process, responsible for the reading a configuration
file, and creating the initial populations. Second, the \texttt{Controller} process,
responsible for the migration between populations, and keeping track of the
iterations of the algorithm. The \texttt{Message Queue} process runs the \texttt{Input} and \texttt{Output}
queues. Finally, there is at least one \texttt{Stateless Function} process responsible
for running the isolated algorithms. In the example, two processes \texttt{PSO} and \texttt{GA},
are shown.

The algorithm starts with the \texttt{Setup} process pulling a configuration message from
the \texttt{Experiments} queue, this is not shown in the diagram. The configuration message 
includes all the parameters needed to execute the algorithm, among other things,
the number of populations, the number of individuals, and the number of iterations
of the algorithm. We give more details about the configuration data structure
in Section \ref{experiment_flow}. % Missing  % added 
Once the configuration is read, we can follow the path of 
messages as follows:

\begin{enumerate}
\item In this step, the specified number of populations are created according to the 
parameters found in the configuration structure. The population at this moment is just
static data, including each individual inside. Each population includes a metadata
section where its algorithm and execution parameters are specified. For instance, 
for a GA, the mutation rate, type of crossover operator, and other values are indicated.

\item Each population is then pushed to the \texttt{Input} queue, so they can be consumed 
by Stateless functions responsible for the execution of the search.  % Why the caps in Stateless? - JJ  
% It refers to the swimlines in the diagram, maybe I should point it out  - Mario
% Added a comment above. - Mario

\item One or more \texttt{Stateless functions} are constantly pulling population messages
from the \texttt{Input} queue. Taking the population data and metadata as parameters, these 
functions are responsible for the actual execution of the optimization algorithm. 
They take the current state of the population and run the algorithm for a certain 
number of iterations.

\item Once they finish the execution, the population state
is packed again along with additional metadata about the execution of the algorithm.
The resulting populations are now pushed to the Output queue. Once finished, another 
population is pulled from the queue.

@@

\item Otherwise, it passes the stream of messages 
to a migration process, where populations are mixed with each other. New populations 
are generated from this migration, and they are again pushed to the Input queue to continue 
in a loop. The \texttt{Controller} is also responsible of logging the metadata received with the
messages.  
\end{enumerate}

@@

Next, we compare the message-driven architecture we presented against four other
proposals for multi-population algorithms found in the literature. We center the
comparison on the coupling and communication between the main components:
populations, processing nodes, and algorithms.  In Figure \ref{fig:population_message} we show the
main components of a message-driven architecture. We can see that the algorithm
and populations are separated, and the \texttt{Population queue} keeps the state
of populations. While algorithms only need the populations as parameters
for their execution.

@@

In contrast, in the classic Island model shown in Figure
\ref{fig:classicisland}, algorithms are coupled with a population, and the same processing
node keeps both the algorithm process and the state of the population.  Other
processing nodes follow the same configuration and form of execution, and they
only pass specific individuals between them.  In this model, adding additional
processing nodes, on-runtime, can be more difficult, because they need to have
an addressing mechanism. % JJ, help I don't know if this is clearly understood, a mean like a DNS - Mario

@@

A typical design pattern used to alleviate the above drawbacks is the use of a
central repository or pool of individuals that is available to all processing
nodes.  In Figure \ref{fig:pool_island} we show the main components of a pool-based
multi-population algorithm. Although algorithms and population state remains
coupled, now, processing nodes do not communicate directly with each other.
Instead, they interchange individuals with the pool. Communication between nodes
is not affected if the system adds or removes nodes because they do not know
about each other.  Since the state of the global population is stored in
external processing nodes, the system needs additional communication and
processing for keeping track of each population.

We show another pool-based approach in Figure  \ref{fig:population_pool}. In
this design, the global population is stored in a central repository, while
isolated algorithms take random samples of the global population and use this
temporal population as parameters. An advantage of this design is that the
sampling provides a type of migration between isolated populations. A problem
found is that when a processing node returns a population to the pool, the
state of the population is lost. But having the algorithm decoupled opens the
possibility of implementing the system using serverless functions.

@@

The EvoSwarm application uses configuration files to run the noiseless BOBB testbed
experiments.  As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The { \em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a defualt value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

To evaluate the effect of these parameters on the scalability of the system, we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function has been used because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.

@@

\begin{table}
    \small
    \caption{Parameters used in the experiments with ten populations
    }
    \label{tab:params:10}
    \vspace{0.25cm}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
      Populations      & 10 & 10 & 10 & 10 & 10  & 10  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

\begin{table}
    \small
    \caption{Parameters used in the experiments that used five populations
    }
    \label{tab:params:5}
    \vspace{0.25cm}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
      Populations      & 5 & 5 & 5 & 5 & 5  & 5  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{sp_size_worker}
    \caption{Time required, in seconds, to complete an instance of the function, with different number of workers and populations sizes }
    \label{fig:spworker}
\end{figure}

@@

We summarize the results of the experiment in Figure \ref{fig:spworker}.  The
experiment consisted of executing the same benchmark on EvoSwarm twice but
changing the number of populations used in each run. The left column has the
results when using five populations, and the right side shows the results for
ten.  Graphics show the execution time achieved by a certain number of workers.
Each box-plot indicates the time required for the execution of 15  function
instances, in seconds.  Each row has a different scale on the y-axis, to
compensate for the dimension increments. It is important to notice that the
execution time can also be affected by the algorithm finding a solution with an
error smaller than the threshold, before completing all the FEs.

As expected, in most cases, increasing the number of workers reduces the time
required for the execution of an instance. Mainly when using the same parameters
(number of populations and number of workers ). But, when we compare the two
options, we find some subtle differences:

When using one or two workers, using 5 populations is better in most cases. A
reason for this could be that the migration of individuals takes more time to
reach the worker. For instance, when a single worker has 10 populations, only
after processing all, it will start receiving mixed populations.  Although, in
all cases, and even a single worker, populations are heterogeneous, this
advantage takes more time to propagate. With four workers, almost the same
results can be achieved regardless of our selection between five or ten
populations. Also, there is no much difference between 4 and 8 workers, when
using five populations. In general, we can observe that keeping a ratio close to
1:1 between the number of workers and populations exploits better the computing
resources available.

@@

>>>>

Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously.
The fitness of each individual can be evaluated independently of 
others, and similarly, each population could evolve in isolation. Since 
earlier works, researchers have been proposing some form of parallelization
\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
The island model was one of the first techniques proposed for parallelization,
which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
The concept was to divide a large population into communicated subpopulations. 
Since then, other population-based algorithms have adopted the concept, 
and researchers have found additional benefits
besides the execution speed; these include avoiding a premature convergence and
maintaining the diversity of the global population \cite{li2015multi}. Researchers use the term
multi-population based methods when generally referring to techniques using
subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
% Comments below - Mario
Moreover, parallel implementations can run in two ways: synchronous and
asynchronous. When operations run in parallel but must maintain synchronism with
the other operations,  all operations need to finish to continue to the next
step. For instance, in a  parent/worker model, the parent needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other. Following the previous example, now the parent operation can
continue to the next iteration, even if a worker has not finished its work.  In
the literature, we can find many asynchronous algorithms
\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time,
and scalability. In the particular case of asynchronous and cloud-native
multi-population algorithms, recent works propose an asynchronous communication
through a central repository \cite{sofea:cec2012, JSON} or message queues
\cite{salza2019speed, guervos2018introducing}, in this work we follow this
practice.

@@

When designing efficient multi-population algorithms, researchers need to
consider additional issues \cite{Ma2019}. These include the number and size of subpopulations,
the interaction between them, the search area of subpopulations, and the search
strategy and parametrization of each subpopulation. We must take into account
these high-level issues when designing a parallel architecture. Researchers 
and algorithm designers willing to  adapt to the cloud their
multi-population solutions face the additional challenges:
\begin{itemize}
    \item They need to change their current solutions to a reactive architecture in which processes
    communicate with each other by exchanging messages asynchronously and react
    to a continuous stream of data. One of the options is to consider
    sub-populations as messages to be modified asynchronously by functions.

@@

Taking these factors into account, in this work, we present EvoSwarm, 
a multi-container Docker application that reactively processes isolated and
heterogeneous subpopulations. The platform uses the Docker Compose tool for
defining and running experiments as multi-container applications. 
% Docker compose is not really for production, more for development environments, 
% you have to justify this. Also, it does not handle scaling so need to justify it too - JJ
% justification: - Mario
We preferred the use of a container-based application for this prototype because
it is more suitable for development and replication without the need for
additional components. Nevertheless, the application is compatible with other
orchestration and deployment technologies if more scalability is needed. 
With additional configuration settings, applications can be deployed 
to a Docker Swarm or a Kubernetes system.

Docker Container images host the stateless functions responsible for running 
a search strategy on subpopulations, these functions could also be ported
directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
% Need to improve the next paragraph or maybe move it to another place - Mario
Researchers then define the configuration of services used by
a multi-population algorithm using a YAML file. Any researcher can deploy and
start all the services from this configuration file with a single command. Once
the services are up, multiple instances of an experiment can be sent to a queue
for their execution. The platform includes containerized services that are
compatible with other container orchestration technologies like Kubernetes and
Docker Swarm. The services included are a Redis Server used for the Message
Queue implementation and logging and an Experiment Controller implemented in
Python. Developers of sub-population processing functions have the freedom to
implement new population-based algorithms or operators in the language they
choose inside a container, as they only need to subscribe to a channel of the
Message Queue and have the ability to read the population and parameters as JSON  
file.

@@

% Also, multipopulation design of what?x
Multi-population based methods divide the original population into
smaller subpopulations or islands, with every subpopulation carrying out the
algorithm independently, with synchronous or asynchronous communication with the
rest of the islands. % exemplify how this is done in different
                     % metaheuristics, especially those we will be
                     % using here - JJ
This relative isolation helps in maintaining an overall
diversity since each subpopulation will search in a particular area, at least
between communications. The recombination mentioned above (mixing) or migration
between subpopulations is needed to avoid a premature convergence of candidate
solutions since smaller populations are known to perform better for a given
problem than bigger populations. % Smaller populations converge
                                % prematurely or perform better? - JJ
However, it gives them the added advantage of
parallel operation. Additionally, and in some cases, multi-population algorithms
scale better than expected due to the interaction between the algorithm and the
parallelism of the operation \cite{ALBA20027}. % Need to clarify this
                                % and if it extends to Swarm
                                % algorithms too
                                % Also, is your objective to achieve this superlinear speedup? - JJ

However, in most cases, algorithms applied to each subpopulation are
homogeneous, or at any rate, the same variant of the algorithm. As long as this
parallel operation is not synchronous, other population-based algorithms, or, as
a matter of fact, any algorithm, could be easily integrated. That is why several
works based on multi-population are heterogeneous, integrating various
optimization algorithms, and often performing better than single-population or
homogeneous optimization algorithms \cite{wu2016differential,nseef2016adaptive}.

Heterogeneous algorithms add another degree of freedom to the problem of finding
the correct parameter settings for an algorithm; because some parameters affect
the accuracy of the solution and the convergence speed of the algorithms as they
tip the balance between exploration and exploitation of the search space. On the
other hand, current studies show that by having a high number of subpopulations
interacting in parallel, the effect of the individual parameters of each
subpopulation is compensated by those selected in other
subpopulations. % reference - JJ
In this
work, we will use random settings within a specific range as results have shown
this is a valid solution to this problem. % This should have been
                                % mentioned in the introduction - JJ

Some parameters, specially the population size, are
kept fixed in order to control more easily the execution of the algorithm. For
instance, by having the size of subpopulations fixed, it is easier to control
the number of evaluations and the communication costs, when the algorithm is in
operation. % And to keep them more or less synchronized, which can be
           % interesting, too... But this should go to the
           % introduction also - JJ

Combining multiple algorithms, with different parameters, interacting with each
other at the same time, can benefit from the strengths of each. For instance, a
genetic algorithm could find a promising global solution that is not optimal
while another algorithm, more suitable for a local search, finds the global
optimum. This approach has been followed extensively in recent years with
success. % References - JJ
Moreover, there is a need for frameworks, architecture, and
implementation models that can allow researchers the development of new
parallel, asynchronous, heterogeneous, and parameter-free algorithms in a scalable way.

@@

The most elemental data structure used throughout the system is the subpopulation; and at a higher
level, we can view the system as a continuous stream of subpopulations
flowing
% We talk about "multi-population" elsewhere, including the
% abstract. We should unify terminology - JJ
% "multi-population" is the whole thing, a subpopulation is a member of that.
% "A multi-population has many subpopulations or should we call them just populations?"
% I will add a comment about this in the Multi-population section -
% Mario
% use the usual "multi-deme" terminology. A population has different
% sub-populations nad uses a multi-deme implementation - JJ
% Let's talk about it, I am using the Multi-population terminology 
% maybe multi-dem is more common in the EA community but not 
% much in the Swarm Intelligence side?  
% Any how, we decide and explain the difference so readers of both 
% camps are Ok with it. 
% I prefer the use of population and something like the global population - Mario

from one process to the other. By "continuous stream," we mean that ideally,
everything must happen asynchronously without components needing to wait for
others. In this kind of system, this streaming functionality is normally
implemented by using a message queue system. For instance, if one process needs
to communicate a subpopulation to another, it must pack the sub-population into
a message, and then send the message 
% I don't know if
                                % asynchronously applies here. It's
                                % simply sent - JJ 
                                % Meaning it does not wait for a
                                % respone, it continues doing other
                                % stuff - Mario
% That's rather "fire and forget". Or simply say it's sent without
% waiting for a response - JJ
% corrected - Mario
to the queuing system without waiting for a response; this
means that after the process pushes the message, it continues its execution
without waiting for a result. On the other end, the recipient subscribes to the
message queue by defining an event handler method that is going to be triggered when a
message is pushed to it. The two components at each end of the queue are
Producers that push messages to the queue and Consumers that pull them. This
pattern is an essential component of a reactive architecture and makes
it 
highly scalable; one of the reasons for this is the use of functions
that have no secondary effects or access global storage, that is,
stateless functions. These functions do not need to read, keep, or modify data outside of the method. If
Consumers are implemented using stateless functions, and thus, there is no difference
between having one or many copies of the same function pulling work from the queue all at the
same time, because no side effects, including unwanted resource
locking problems,  are going to result from this concurrency.
% This is important and we should probably mention that in the
% introduction - JJ 
Based on these general principles we define next the proposed model.

@@

\subsection{Event Driven Model} 
\label{edm}
Based on a reactive architecture we proposed in a previous work
\cite{guervos2018introducing}, we now describe the general architecture shown in
Figure \ref{fig:kafkEO}. Again, we can explain the model using the analogy
of producers and consumers of messages. First, we can see that
there are two queues, one labeled \texttt{Input}, and the other one \texttt{Output}. In the diagram, 
push operations on a queue are represented by solid arrows connecting to the left side
of the queue box, and pull or pop operations as solid arrows leaving from the right side.
Also, the architecture has at least four processes indicated in the diagram as
swimlanes: First, the \texttt{Setup} process, responsible for the reading a configuration
file, and creating the initial subpopulations. Second, the \texttt{Controller} process,
responsible for the migration between subpopulations, and keeping track of the
iterations of the algorithm. The \texttt{Message Queue} process runs the \texttt{Input} and \texttt{Output}
queues. Finally, there is at least one \texttt{Stateless Function} process responsible
for running the isolated algorithms. In the example, two processes \texttt{PSO} and \texttt{GA},
are shown.

The algorithm starts with the \texttt{Setup} process pulling a configuration message from
the \texttt{Experiments} queue, this is not shown in the diagram. The configuration message 
includes all the parameters needed to execute the algorithm, among other things,
the number of subpopulations, the number of individuals, and the number of iterations
of the algorithm. We give more details about the configuration data structure
in Section \ref{experiment_flow}. % Missing  % added 
Once the configuration is read, we can follow the path of 
messages as follows:

\begin{enumerate}
\item In this step, the specified number of populations are created according to the 
parameters found in the configuration structure. The population at this moment is just
static data, including each individual inside. Each subpopulation includes a metadata
section where its algorithm and execution parameters are specified. For instance, 
for a GA, the mutation rate, type of crossover operator, and other values are indicated.

\item Each subpopulation is then pushed to the \texttt{Input} queue, so they can be consumed 
by Stateless functions responsible for the execution of the search.  % Why the caps in Stateless? - JJ  
% It refers to the swimlines in the diagram, maybe I should point it out  - Mario
% Added a comment above. - Mario

\item One or more \texttt{Stateless functions} are constantly pulling subpopulation messages
from the \texttt{Input} queue. Taking the population data and metadata as parameters, these 
functions are responsible for the actual execution of the optimization algorithm. 
They take the current state of the population and run the algorithm for a certain 
number of iterations.

\item Once they finish the execution, the population state
is packed again along with additional metadata about the execution of the algorithm.
The resulting populations are now pushed to the Output queue. Once finished, another 
subpopulation is pulled from the queue.

@@

\item Otherwise, it passes the stream of messages 
to a migration process, where subpopulations are mixed with each other. New populations 
are generated from this migration, and they are again pushed to the Input queue to continue 
in a loop. The \texttt{Controller} is also responsible of logging the metadata received with the
messages.  
\end{enumerate}

@@

Next, we compare the message-driven architecture we presented against four other
proposals for multi-population algorithms found in the literature. We center the
comparison on the coupling and communication between the main components:
subpopulations, processing nodes, and algorithms.  In Figure \ref{fig:population_message} we show the
main components of a message-driven architecture. We can see that the algorithm
and subpopulations are separated, and the \texttt{Population queue} keeps the state
of subpopulations. While algorithms only need the populations as parameters
for their execution.

@@

In contrast, in the classic Island model shown in Figure
\ref{fig:classicisland}, algorithms are coupled with a subpopulation, and the same processing
node keeps both the algorithm process and the state of the subpopulation.  Other
processing nodes follow the same configuration and form of execution, and they
only pass specific individuals between them.  In this model, adding additional
processing nodes, on-runtime, can be more difficult, because they need to have
an addressing mechanism. % JJ, help I don't know if this is clearly understood, a mean like a DNS - Mario

@@

A typical design pattern used to alleviate the above drawbacks is the use of a
central repository or pool of individuals that is available to all processing
nodes.  In Figure \ref{fig:pool_island} we show the main components of a pool-based
multi-population algorithm. Although algorithms and subpopulation state remains
coupled, now, processing nodes do not communicate directly with each other.
Instead, they interchange individuals with the pool. Communication between nodes
is not affected if the system adds or removes nodes because they do not know
about each other.  Since the state of the global population is stored in
external processing nodes, the system needs additional communication and
processing for keeping track of each subpopulation.

We show another pool-based approach in Figure  \ref{fig:population_pool}. In
this design, the global population is stored in a central repository, while
isolated algorithms take random samples of the global population and use this
temporal subpopulation as parameters. An advantage of this design is that the
sampling provides a type of migration between isolated subpopulations. A problem
found is that when a processing node returns a subpopulation to the pool, the
state of the subpopulation is lost. But having the algorithm decoupled opens the
possibility of implementing the system using serverless functions.

@@

The EvoSwarm application uses configuration files to run the noiseless BOBB testbed
experiments.  As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The { \em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a defualt value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of  Function Evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

To evaluate the effect of these parameters on the scalability of the system, we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.

@@

\begin{table}
    \small
    \caption{10 Populations experiment parameters
    }
    \label{tab:params:10}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
      Populations      & 10 & 10 & 10 & 10 & 10  & 10  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

\begin{table}
    \small
    \caption{5 Populations experiment parameters
    }
    \label{tab:params:5}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
      Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
      Populations      & 5 & 5 & 5 & 5 & 5  & 5  \\ \hline
      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
    \end{tabular}
\end{table}

@@

@@

\begin{figure}[h!tb]
  % Is the axis correct? Seconds per instance or total time? Please
  % change that - JJ
    \centering
    \includegraphics[width=\textwidth]{sp_size_worker}
    \caption{Time required to complete experiments with different number of workers and populations sizes }
    \label{fig:spworker}
\end{figure}

@@

\begin{figure*}[h!tb]
    \begin{tabular}
        {c@{\hspace*{-0.00001\textwidth}}
         c@{\hspace*{-0.00001\textwidth}}
         c@{\hspace*{-0.00001\textwidth}}
        }
    GA  &  PSO & GA \& PSO\\   
    \includegraphics[width=0.33\textwidth]{GAOnly_f001}&
    \includegraphics[width=0.33\textwidth]{PSOOnly_f001}&
    \includegraphics[width=0.33\textwidth]{GAPSO_f001}\\

    \includegraphics[width=0.33\textwidth]{GAOnly_f002}&
    \includegraphics[width=0.33\textwidth]{PSOOnly_f002}&
    \includegraphics[width=0.33\textwidth]{GAPSO_f002}\\

    \includegraphics[width=0.33\textwidth]{GAOnly_f003}&
    \includegraphics[width=0.33\textwidth]{PSOOnly_f003}&
    \includegraphics[width=0.33\textwidth]{GAPSO_f003}\\

    \includegraphics[width=0.33\textwidth]{GAOnly_f004}&
    \includegraphics[width=0.33\textwidth]{PSOOnly_f004}&
    \includegraphics[width=0.33\textwidth]{GAPSO_f004}\\

    \includegraphics[width=0.33\textwidth]{GAOnly_f005}&
    \includegraphics[width=0.33\textwidth]{PSOOnly_f005}&
    \includegraphics[width=0.33\textwidth]{GAPSO_f005}\\
    \end{tabular}
    \vspace{-3ex}
     \caption{test}
    \end{figure*}

    \begin{figure}[h!tb]
        \begin{tabular}
            {c@{\hspace*{-0.00001\textwidth}}
             c@{\hspace*{-0.00001\textwidth}}
            }

        \includegraphics[width=0.45\textwidth]{ppfigs_f001}&
        \includegraphics[width=0.45\textwidth]{ppfigs_f002}\\

        \includegraphics[width=0.45\textwidth]{ppfigs_f003}&
        \includegraphics[width=0.45\textwidth]{ppfigs_f004}\\
